---
title: "Homework 3"
author: "Charlotte Fowler"
date: "10/9/2019"
output: github_document
---

```{r}
#load packages used
library(tidyverse)
library(p8105.datasets)
```


# Problem 1 

```{r}
#load data
data("instacart")
```



This dataset has `r nrow(instacart)` rows and `r ncol(instacart)` variables. Each row represents a different item ordered by a instacart customer. 
We have information on the order, such as `order_id`, `order_hour_of_day`, `order_dow` or the day of week in which the order occured, and `add_to_cart_order` which shows the order in which items were added to the cart. 
We also have information on the customer such as `user_id`, `order_number` or the number of prior orders (+1) this customer has had, and `days_since_prior_order`. 
Lastly we have information on the product being requested such as `product_id`, `product_name`, `aisle_id`, `aisle`, `department_id`,  and `department`. 

For example, we know that customer 127134, in their 10th order from instacart, on a Friday at 9am added 'Hash Brown Potato Patties', in frozen appetizers sides aisle from the frozen department, for the first time as the 21st item in their cart. We can also see that 19 other orders in the data contained 'Hash Brown Potato Patties'.  

```{r results=FALSE}
#searching for Hash Brown Potato Patties in dataset
instacart %>% 
  filter(product_name == 'Hash Brown Potato Patties') %>% 
  nrow()
```

```{r}
#counting number of aisles
instacart %>% 
  group_by(aisle) %>% 
  count() %>% 
  nrow() 
```


```{r}
#finding top four most popular aisles 
instacart %>% 
  group_by(aisle) %>% 
  count() %>% 
  arrange(-n) %>% 
  .[0:4,] %>% 
  knitr::kable()
```






There are 134 aisles. The most popular are fresh vegetables, fresh fruits,  packaged vegetables fruits, and yogurt. 



```{r}
#Building bar chart of most popular aisles 
instacart %>% 
  group_by(aisle) %>% 
  count() %>% 
  filter(n>10000) %>% 
  inner_join (instacart, by = 'aisle') %>% 
  select(aisle, n, department) %>% 
  distinct() %>% 
  arrange(-n) %>% 
  ggplot(aes(x = reorder(aisle, -n), y = n, fill = department)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Most Popular Aisles", y = "Number of Orders", x = "Aisle")

```


Produce aisles appear the most popular, followed by some dairy and eggs aisles. There are a few aisles such as fresh vegetables and fruits which have far more orders than any others.  


```{r}
#picking top 3 products from baking aisle
baking_aisle = instacart %>% 
  filter(aisle == "baking ingredients") %>%
  select(product_name, aisle) %>%   
  group_by(aisle, product_name) %>% 
  count() %>% 
  arrange(-n) %>%
  .[0:3,]

#picking top 3 products from dog aisle
dog_aisle = instacart %>% 
  filter(aisle == "dog food care") %>%
  select(product_name, aisle) %>%   
  group_by(aisle, product_name) %>% 
  count() %>% 
  arrange(-n) %>%
  .[0:3,]

#picking top 3 products from veg aisle
veg_aisle = instacart %>% 
  filter(aisle == "packaged vegetables fruits") %>%
  select(product_name, aisle) %>%   
  group_by(aisle, product_name) %>% 
  count() %>% 
  arrange(-n) %>%
  .[0:3,]
```

```{r}
#binding 3 tables together and cleaning
three_aisles = 
  bind_rows(baking_aisle, dog_aisle, veg_aisle) %>% 
  rename(
    "Top Products" = product_name, 
    "Aisle" = aisle,
    "Number of Orders" = n
    ) 

knitr::kable(three_aisles)
```

The packaged vegetables fruits aisle's top products have far more orders than the other two. The top dog food care products still have relatively few orders. 













```{r}
#making table of apples and ice cream and time ordered
instacart %>% 
  filter(product_name %in% c('Pink Lady Apples', 'Coffee Ice Cream')) %>% 
  select(order_hour_of_day, order_dow, product_name) %>% 
  group_by(order_dow, product_name) %>% 
  summarize(mean_hour = round(mean(order_hour_of_day), digits = 2)) %>% 
  pivot_wider(names_from = product_name, values_from = 'mean_hour') %>% 
  ungroup() %>% 
  rename(Day = order_dow,
         'Pink Lady Apple: mean hour' ='Pink Lady Apples',
         'Coffee Ice Cream: mean hour' = 'Coffee Ice Cream'
         ) %>% 
  mutate(
    Day = recode(Day, '0' = "Monday", '1' = "Tuesday", '2'="Wednesday", '3'="Thursday", '4'="Friday", '5'="Saturday", '6'="Sunday"), 
  ) %>% 
  knitr::kable() 
```


People appear to order their ice cream later in the day than apples. In general the apples are typically ordered around 11 am, and the ice cream is ordered around 2 pm. Orders on the weekend appear to be earlier than orders on the weekdays. 

# Problem 2 




```{r} 
#reading in data
data("brfss_smart2010")


#data cleaning for brfss
brfss_smart2010 = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  separate(locationdesc, 
           into = c('state', 'county'),
           sep = " - ") %>% 
  select(-locationabbr) %>% 
  filter(
    topic == 'Overall Health',
    response %in%  c('Excellent', 'Good', 'Very good', 'Fair', 'Poor')
  ) %>% 
  mutate(response = forcats::fct_relevel(response, c("Poor", "Fair", "Good", "Very good", "Excellent")))
```











```{r}
#calculating states with 7+ locations in 2002
brfss_smart2010 %>% 
  filter(
    year==2002
  ) %>% 
  select(state, county) %>% 
  distinct() %>% 
  count(state) %>% 
  filter(n>=7) %>% 
  knitr::kable() 
```

```{r}
#calculating states with 7+ locations in 2010
brfss_smart2010 %>% 
  filter(
    year==2010
  ) %>% 
  select(state, county) %>% 
  distinct() %>% 
  count(state) %>% 
  filter(n>=7) %>% 
  knitr::kable() 
```


In 2002, Conneticut, Florida, Massachusetts, North Carolina, New Jersey and Pennsylvania were observed at 7 or more locations. In 2010, California, Colorado, Florida, Massachusetts, Maryland, North Carolina, Nebraska, New Jersey, New York, Ohio, Pennsylvania, South Carolina, Texas and Washington were observed in 7 or more locations.  




```{r warning = FALSE}
#creating new data set with mean values per year, and making spaghetti plot
brfss_smart2010 %>% 
  filter(response == 'Excellent') %>% 
  group_by(state, year) %>% 
  summarize(mean_value = mean(data_value)) %>% 
  ggplot(
    aes(x = year, y = mean_value, color = state)) +
  geom_line() + 
  labs(x = 'Year', y = 'Mean Data Value', title = 'Data by State across Years') + 
  viridis::scale_color_viridis(
    discrete = TRUE, 
    name = "State"
    ) +
  theme_bw() 
```




Some states appear to have much higher mean values compared to others. Additionally, some states are very fluctual, while others have relatively stable mean values across time. In most states the mean value appears to have risen around 2006, and fallen shortly after. 

```{r}
#making 2 panel bar graph 
brfss_smart2010 %>% 
  filter(
    year %in% c(2006, 2010), 
    state == 'NY'
  ) %>% 
  ggplot(aes( x = response, y = data_value)) + 
  geom_bar(stat = "identity") + 
  facet_grid(~year) + 
  labs(title = "New York 2006 and 2010 Responses", x = "Response", y = "Data Value")
  
```


The values appear higher in all categories in 2010 compared to 2006. In both years, there are higher values on the positive versus negative responses. Very good in both years was the highest response. 


# Problem 3 



```{r}
#tidying the data
accel_data = read_csv("./data/accel_data.csv")
accel_data = accel_data %>% 
  pivot_longer(
    'activity.1':'activity.1440',
    values_to = 'activity', 
    names_to = 'minute'
  ) %>% 
  separate(minute, 
           into = c('non', 'minute'),
           sep = 9) %>% 
  select(-non) %>% 
  mutate(
    weekend = ifelse(day == c("Sunday", "Saturday"), TRUE, FALSE ),
    minute = as.numeric(minute),
    day = factor(day),
    week = factor(week)
  )
```

The dataset `accel_data` consists of `r nrow(accel_data)` observations by `r ncol(accel_data)` variables. Each observation represents a minute of data observed by the accelerometer. We have information on the `minute`, `day`, and `week`, as well as a logical variable on whether it was the `weekend`. We also have information on `day.id`, or how many days the patient has been wearing the device. Lastly, we have the `activity` count for that minute.






```{r}
#looking at sum for each day
accel_data %>% 
  group_by(day_id) %>% 
  summarize(activity_sum =(sum(activity))) 
```
It appears that on weekends the individual was either occasionally much less active, or would not wear the device. For example on Saturday, the 24, and Saturday the 31st, the there is practically no activity. Activity appears much more consistent during the week, and more variable on weekends. 




```{r}
#making graoh with activity across each day
accel_data %>% 
  ggplot(aes(x = minute, y = activity, group = day_id, color = day)) +
  geom_line(size = 0.5) +
  viridis::scale_color_viridis(
    discrete = TRUE, 
    name = "Day"
    ) + 
  labs(title = "Activity by Day", x = "Minute of Day", y = "Activity")
```

The most active period of the day is aroun minute 1250, or around 8 PM. The individual was least active from about 10 PM to 3 AM. There were occasional peaks of activity around lunchtime and 4 PM. The inidividual appears slightly more active on the weekend versus weekdays. 
